Supplmental Information : Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-10-09 Ben_Manuscripts/structure_paper/figures/xrd.tex
Rethinking the scope of the paper. Renamed folders to be more descriptive

\documentclass{article}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath} % or simply amstext
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}
\newcommand{\angstrom}{\textup{\AA}}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{float}
\title{Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes}
\author{Benjamin J. Coscia \and Douglas L. Gin \and Richard D. Noble \and Joe Yelk \and Matthew Glaser \and Xunda Feng \and Michael R. Shirts} 
\begin{document}
  \bibliographystyle{ieeetr}
  \graphicspath{{./figures/}}
  \maketitle
  \section*{Introduction}

  Nanostructured membrane materials have become increasingly popular for 
  aqueous separations applications such as desalination and biorefinement
  because they offer the ability to control membrane architecture at the
  atomic scale allowing the design of solute-specific separation membranes. \cite{humplik_nanostructured_2011}
  \begin{itemize}
    \item Most membrane-based aqueous separations of small molecules can 
    be achieved using reverse osmosis (RO) or nanofiltration (NF) \cite{van_der_bruggen_review_2003}
  \end{itemize}

  While RO and NF have seen many advances in the past few decades, they 
  are far from perfect separation technologies.
  \begin{itemize}
    \item \textit{RO membranes}
    \item Inconsistent performance : Current state-of-the-art RO membranes are unstructured with
    tortuous and polydisperse diffusion pathways which leads to 
    inconsistent performance \cite{song_nano_2011}
    \item High energy requirements : Necessarily high feed pressures 
    drive up energy requirements which strains developing regions and
    contributes strongly to CO\textsubscript{2} emissions. \cite{mcginnis_global_2008}
    \item Separation based on differences in solubility and diffusivity:
    Moreover, designing RO membranes to achieve targeted separations of 
    specific solutes is nearly impossible because various solutes dissolve
    into and diffuse through the polymer matrix at different rates. \cite{wijmans_solution-diffusion_1995}
    \item At best, one can exploit these differences to create a functional
    selective barrier.
    \item \textit{NF membranes}
    \item NF was introduced as an intermediate between RO and ultrafiltration,
    having the ability to separate organic matter and salts on the order of 
    one nanometer in size.
    \item Larger and well-defined pores drive down energy requirements while
    still affording separation of solutes as small as ions to some degree \cite{van_der_bruggen_review_2003}
    \item NF is often used as a precursor to reverse osmosis
    \item Unfortunately, NF membranes, like RO, possess a pore size 
    distribution which limits their ability to perform precise separations \cite{bowen_modelling_2002}
  \end{itemize}

  Nanostructured membranes can bypass many of the performance issues which
  plague traditional NF and RO membranes.
  \begin{itemize}
    \item Tune size and functionality of building blocks to control pore
    size and shape: One can accomplish targeted separations with high 
    selectivity by tuning shape, size and functionality of the molecular
    building blocks which form these materials. % BJC: "these materials" --> "nanostructured membranes", or is that redundant?
    \item As a result, solute rejecting pores can have their sizes tuned
    uniformly, resulting in strict size cut-offs.
    \item Entirely different mechanisms may govern transport in a given
    nanostructured material which can inspire novel separation techniques.
  \end{itemize}

  Development of nanostructured materials has been limited by the ability
  to synthesize and scale various fundamentally sound technologies.
  \begin{itemize}
    \item Graphene sheets are atomically thick which results in excellent
    permeability but defects during manufacturing severely impact 
    selectivity. \cite{cohen-tanugi_multilayer_2016}
    \item Molecular dynamics simulations of carbon nanotubes show
    promise \cite{humplik_nanostructured_2011} but synthetic techniques are 
    unable to achieve scalable alignment and pore monodispersity.\cite{hata_water-assisted_2004,maruyama_growth_2005}
    \item Zeolites have sub-nm pores with a narrow pore size 
    distribution and MD simulations exhibit complete rejection of solvated ions, \cite{murad_molecular_1998}
    however, experimental rejection was low and attributed to interstitial
    defects formed during membrane synthesis \cite{li_desalination_2004}
    \item There is a need for a scalable nanostructured membrane
  \end{itemize}

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. 
  \begin{itemize}
    \item LLCs share the characteristic ability of nanostructured membrane
    materials to create highly ordered structures with the added benefits
    of low cost and synthetic techniques feasible for 
    large scale production \cite{feng_scalable_2014}
    \item Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
    phase. The presence of water results in the H\textsubscript{II} phase.
    \item In both cases, monomers assemble into mesophases made of hexagonally
    packed, uniform size, cylinders with hydrophilic groups oriented inward
    towards the pore center and hydrophobic groups facing outward.
    \item H\textsubscript{II} and Col\textsubscript{h} phase systems created by
    the monomer named Na-GA3C11 has been extensively studied experimentally \cite{smith_ordered_1997, %BJC: IUPAC chemical name here?
    zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}. 
    \item Until recently, mesophases formed by Na-GA3C11 could not be macroscopically
    aligned, resulting in a low flux membrane, slowing research in the field.
    \item In 2014, Feng et al. showed that the mesophases could be aligned 
    using a magnetic field with subsequent crosslinking to lock the structure
    in place \cite{feng_scalable_2014}
    \item In 2016, Feng et al. showed that the same result could be obtained 
    using a technique termed soft confinement \cite{feng_thin_2016}.
    \item Following this breakthrough, research into LLC membranes has been
    reinvigorated
  \end{itemize}

  A molecular level understanding of LLC membrane structure, enabled by molecular
  dynamics simulations, will provide guidelines to reduce the large chemical space
  available to design monomers for creation of separation-specific membranes.
  \begin{itemize}
%     \item LLCs are versatile and controllable with a \textbf{large chemical design
%     space} available for membrane design  %BJC: redundant
    \item Over the past 20 years, H\textsubscript{II} phase LLC membrane studies have been limited primarily
    to Na-GA3C11 with some characterization done after minor structural modifications
    \cite{resel_structural_2000}.
    \item Rejection studies show that this membrane can not perform separations of
    solutes less than 1.2 nm in diameter because the pores are too large \cite{zhou_supported_2005}.
    \item  We do not yet understand how to reduce the effective pore size or 
    how to tune the chemical environment in the nanopores for effective water
    desalination or small organic molecule separations.
    \item It will be challenging to efficiently narrow down the large design space in 
    a laboratory setting without a robust model.
    \item The only source of predictive modeling has been macroscopic models 
    which likely do not adequately describe transport at these length scales. % BJC: Reference. I think it is just w.r.t. bicontinuous cubic
    \item Choice of head group may play a role in the rejection of charged or uncharged solutes.
    \item Choice of counterion may influence the establishment of a Donnan potential
    affecting the degree to which the membrane can exclude charged species.
    \item A good molecular model should incorporate a detailed picture of the nanoscopic pore 
    structure which will be crucial to understanding the role of monomer structure in 
    membrane design.
    \item Molecular dynamics simulations will provide the required level of detail
  \end{itemize}

  Our approach to constructing a general model will follow the development of a
  model of a specific LLC membrane with sufficient experimental characterization.
    \begin{itemize}
	    \item We have chosen to focus on assemblies formed by Na-GA3C11
	    \item We have also narrowed our scope to the development of 
	    a model of the Col\textsubscript{h} phase membrane.
	    \item Compared to the H\textsubscript{II} phase, the Col\textsubscript{h}
	    phase is a simpler starting point, due to the absence of water, and has
	    more detailed experimental structural data.  % BJC: More detailed because of WAXS. HII doesn't have any (that I know of) WAXS
	    \item Despite having structural data, there is still information which 
	    experiment cannot definitively answer.
    \end{itemize}

%BJC: the following just seems like filler now that I've reorganized
%   A clear picture of the nanoscopic LLC membrane structure, gained by building 
%   a molecular model will provide evidence to answer existing and newly proposed
%   questions.
%     \begin{itemize}
%     \item Despite having structural data, there is still information which 
%     experiment cannot definitively answer
%     \end{itemize}

%MRS: break this down into more hierarchy - there are 11 separate points.  Can they be grouped at all for clearer understanding? elucidate what the aspects are we need a clearer picture. 
%BJC: I moved the logic for choice of monomer to system to model to a separate paragraph (above this). I then moved the other three things I talk about into their own paragraphs to keep the ideas separate (following paragraphs).

%MRS: since the questions we answer in the paper have a lot to do with the structure of the membranes, somewhere relatively early in the paper we need to lay out the facts that 1) these membranes could be really good (I think you do establish this) 2) we really would need to know the structure better to really do useful things with these, and 3) MD is a very helpful way to gain more information about the structure, because of the limited range of experiments we can do with these systems. That is the main rationale, and it should be made clear fairly early in the introduction.  Very important to make your story absolutely as clear as possible.
%BJC: I moved and edited a paragraph that was originally later on to be 2 above this one. I think it does a better job addressing points 2 and 3. And now it addresses those points earlier in the intro (as early as it could go where I think it still makes sense in context)

  Despite having structural data, there is still information which 
  experiment cannot definitively answer. There are several key 
  questions that we intend to answer which will be laid out and numbered
  in subsequent paragraphs.

%BJC: How question (1) and (2) used to be ordered:
%
%  (1) Do monomers partition into and persist as defined monomer layers?
%  H\textsubscript{II} and Col\textsubscript{h} phase LC membrane pores are 
%  thought to be composed of monomer layers stacked on top of each other. 
%  \begin{itemize}
%  	\item Their existence is supported by experimental evidence of strong 
%	$\pi$-$\pi$ stacking interactions in the direction perpendicular to the
%	membrane plane.
%	\item $\pi$-$\pi$ stacking will only occur between monomer head groups which
%	leaves no description of what is happening in monomer tail region
%	\item It is possible that $\pi$-$\pi$ stacking occurs vertically 
%	with no order in reference to neighboring stacked columns of the same pore.  % BJC: not sure if this is actually true
%  \end{itemize}
%
%  (2) If layers do exist, how many monomers constitute a single layer?
%  There has been no definitive answer to the question in literature.
%  \begin{itemize}
%    \item A simple molecular simulation study of a similar molecule suggested
%    that there are 4 monomers in each layer. Their estimation is based on a
%    simulated system containing only 16 total monomers which likely does not sufficiently
%    model the chemical environment present in the real system.~\cite{zhu_methacrylated_2006}. 
%    \item A separate calculation based on the volume of the liquid crystal monomers proposes
%    that there are seven monomers in each layer~\cite{resel_structural_2000}. 
%    \item A molecular model orders of magnitude larger than any other reported atomistic 
%    liquid crystal membrane simulations has the best chance of directly answering this question.
%  \end{itemize}

  Monomers in the Col\textsubscript{h} system are theorized to be partitioned
  into stacked layers which form columnar pores. There has been no definitive
  answer in literature regarding the number of monomers in each layer. We want to
  know (1) If layers do exist, how many monomers constitute a single layer?
  \begin{itemize}
    \item A simple molecular simulation study of a similar molecule suggested
    that there are 4 monomers in each layer. Their estimation is based on a
    simulated system containing only 16 total monomers which likely does not sufficiently
    model the chemical environment present in the real system.~\cite{zhu_methacrylated_2006}. 
    \item A separate calculation based on the volume of the liquid crystal monomers proposes
    that there are seven monomers in each layer~\cite{resel_structural_2000}. 
    \item A molecular model orders of magnitude larger than any other reported atomistic 
    liquid crystal membrane simulations has the best chance of directly answering this question.
    \item We can directly change the layer composition and note its effect on membrane structure.
  \end{itemize}

  (2) Does our model support the existence of layers and if so, how well defined
  are the layers?
  \begin{itemize}
       \item Experimentally, their existence is supported by evidence of strong 
       $\pi$-$\pi$ stacking interactions in the direction perpendicular to the
       membrane plane.
       \item $\pi$-$\pi$ stacking will only occur between the aromatic monomer head groups which
       leaves no description of what is happening in the monomer tail region
       \item The tails may entangle isotropically while stacking order is maintained
       among headgroups. 
  \end{itemize}

   (3) How do monomers in each layer position themselves with respect to surrounding 
   layers?
%     Even if we knew the number of monomers in each layer, we still would not know how 
%     monomers in each layer are positioned with respect to other layers. 
   \begin{itemize}
      \item The $\pi$-$\pi$ stacking interactions may be a driving force of self assembly 
      in this system\cite{gazit_possible_2002} %is thought to be $\pi$-$\pi$ stacking interactions between aromatic headgroups \cite{gazit_possible_2002}. 
      \item Gas phase ab initio studies of benzene dimers have shown a clear energetic
      advantage for parallel displaced and T-shaped $\pi$-$\pi$ stacking conformations versus a
      sandwiched conformation ~\cite{sinnokrot_estimates_2002}.
      \item Substituted benzene rings exhibit an even stronger $\pi$-$\pi$ stacking 
      attraction which favors the parallel displaced configuration in all cases
      except where the substitutions are extremely electron withdrawing
      \cite{waller_hybrid_2006,ringer_effect_2006}.
      \item We can use simulated X-ray diffration patterns to compare the two 
      stacking configurations. 
   \end{itemize}

   (4) Is it necessary to include any water in order to appropriately model the 
   Col\testsubscript{h} phase?
   \begin{itemize}
	\item While the Col\textsubscript{h} phase is described as dry, it is 
	possible that small amounts of ambient water may be leached into the system.
	\item The hydrogen bonding network formed by the water may play a role 
        in structuring the pore.
	\item We can use simulated X-ray diffraction patterns to see if there is
        any meaningful structural difference between a dry and "wet" system
   \end{itemize}

%     \item While we might be able to provide answers to these questions using a 
%     molecular model, there remains the possibility that there is more than one 
%     metastable state associated with a given LLC system.

%  (4) Can the system exist in other metastable states or phases that are not
%  accessed during experiments?
%  There remains the possibility that there is more than one metastable state 
%  associated with a given LLC system.
%  \begin{itemize}
%    \item Simulating a membrane atomistically will require many atoms which further
%    limits the timescales acessible with MD
%    \item It is reasonable to expect that we will generate configurations which 
%    are kinetically trapped in a metastable free energy basin
%    \item We must be able to identify which state is produced experimentally and 
%    why others are not.
%  \end{itemize}

  %Once we have addressed all of the above questions, we must show that the 
  %developed molecular model is consistent with physical observations so that we
  %can rely on conclusions drawn about structural features characteristic of 
  %the system.
  In this study, we build a significantly more realistic atomistic model of LLC membranes
  than has ever previously been done, and explore what new structural information can be gained
  and what structure hypotheses are supported by this model.

  \begin{itemize}
    \item We validate the model using as much experimental information as possible.
    \item We are most interested in reproducing the conclusions about structure
    which have been made from X-ray diffraction (XRD) experiments and in matching ionic
    conductivity measurements~\cite{feng_thin_2016}.
    \begin{itemize}
	    \item We have compared simulated X-ray diffraction patterns to experiment in 
	    order to match major features present in the 2D patterns.
	    \item We calculated ionic conductivity using two agreeing methods.
	    \item We examined the the influence of crosslinking on membrane structure.
    \end{itemize}
    \item The structure-building approach and analysis used in this paper can be readily extended
    to the H\textsubscript{II} phase and other similar LC systems.
  \end{itemize}

  \section*{Methods}

  Liquid crystal monomers were parameterized using the Generalized AMBER Forcefield
  \cite{wang_development_2004} with the Antechamber package \cite{wang_automatic_2006}
  provided with AmberTools16 \cite{case_ambertools16_2016}. Atomic charges were
  assigned using tools from Openeye Scientific Software. All molecular dynamics 
  simulations were run using the latest version of Gromacs 2016. 
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}

  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a
  parameterized monomer.
  \begin{itemize}
    \item Monomers were cooled from 1000K to 50K over 10 nanoseconds.
    \item A low energy configuration was randomly pulled from the trajectory
    and charges were reassigned using the am1bccsym method of molcharge
    shipped with Openeye Scientific software's QUACPAC % BJC: how to cite?
    \item Using the new charges, the monomer system was annealed again and monomer
    configurations were pulled from the trajectory to be used for full
    system construction (Figure~\ref{fig:python}a).
  \end{itemize}

  The timescale for self assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic
  simulation, calling for a more efficient way to build the system. 
  \begin{itemize}
    \item Previous work has shown a coarse grain model self assemble into
    the H\textsubscript{II} phase configuration in $\approx$ 1000 ns \cite{mondal_self-assembly_2013}.
    \item We attempted atomistic self-assembly by packing monomers into a box 
    using Packmol \cite{martinez_packmol:_2009}.
    \item Simulations of greater than 100 ns show no indicators of progress 
    towards an ordered system.  %BJC: include more details in supplment?
    \item To bypass the slow self-assembly process, python scripts are used
    to assemble monomers into a structure close to one of a number of 
    hypothesized equilibrium configurations (Figure~\ref{fig:python}).
  \end{itemize}

  \begin{figure}[!ht]
	\centering
	\includegraphics[width=0.75\linewidth]{build.PNG} %BJC: put an xyz axis with the unit cell
	\caption{(a) A single monomer was parameterized and annealed to produce a low energy
		configuration. (b) Monomers are rotated and assembled into layers with 
		hydrophlic centers. (c) Twenty layers are stacked on top of each other to create
		a pore. (d) Pores are duplicated and placed into a monoclinic unit cell}\label{fig:python}
  \end{figure}

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended 
  periodically.
  \begin{itemize}
    \item Each pore is made of twenty stacked monomer layers with periodic 
    continuity in the z direction, avoiding any edge effects and creating an 
    infinite length pore ideal for studying transport.
    \item A small number of layers is preferred in order to reduce computational
    cost and to allow us to look at longer timescales.
    \item Ultimately, we chose to build a system with 20 monomer layers in each pore
    in order to obtain sufficient resolution when simulating X-ray diffraction patterns.
    This point will be explained in more detail later.
    %MRS: you should give the reasons in the same place you describe the choice, but the supporting data would probably be better in supporting information
    %BJC: so refer to supporting info here? Then put plots of diffraction data with different sized system there. I did not explain the resolution point here because I haven't talked about simulated X-ray diffraction yet and it wouldn't make sense. And it doesn't make sense to talk about XRD this early in the methods but I need to justify the number of layers.
    \item We chose initial guesses for the remaining structural parameters based on
    experimental data and treated them as variables during model development.
  \end{itemize}

  We used experimental wide angle X-ray scattering (WAXS) data (produced as described 
  in~\cite{feng_scalable_2014}) and small angle X-ray scattering (SAXS) data from~\cite{feng_thin_2016}
  to inform our choices of some initial structural parameters (Figure~\ref{fig:SWAXS}). We rely primarily
  on the 2D WAXS data since it encodes all structural details down to the sub-nm scale.
  \begin{itemize}
        \item There are five major features of interest present in the 2D experimental
        pattern shown in Figure~\ref{fig:WAXS}.
        \item The first is located at $q_z$ = 1.7 \angstrom$^-1$,
        corresponding to a real spacing of 3.7 \angstrom~. The reflection is
        attributed to $\pi$-$\pi$ stacking between aromatic rings in the direction
        perpendicular to the membrane plane, or z-axis \cite{feng_scalable_2014}. For simplicity, this
        reflection will be referred to as R-$\pi$.
        \item A weak intensity line is located at exactly half the $q_z$ value of
        R-$\pi$ ($q_z$ = 0.85 \angstrom$^-1$), corresponding to a
        real space periodic spacing of 7.4 \angstrom~. This reflection has been
        interpreted  as 2\textsubscript{1} helical ordering of aromatic rings
        along the z axis meaning if the positions of the aromatic rings can
        be traced by a helix, then for each turn in the helix, there should be
        two aromatic rings. For this reason it will be referred to as R-helix.
        \item A third major reflection is marked by a low intensity ring located
        at r = 1.4 \angstrom$^-1$. The real space separation
        corresponds to 4.5 \angstrom~ which is characteristic of the average
        spacing between packed alkane chains. This reflection will be called R-alkanes.
        \item Within R-alkanes, are four spots of higher relative intensity which
        will be called R-spots. All are located $\approx 40$ degrees from the $q_z$ axis
        in their respective quadrants. In many liquid crystal systems this can be
        explained by the tilt angle of the alkane chains with respect to the xy plane. % BJC: References
        \item The first corresponds to the spacing and symmetry of
        the d\textsubscript{100} plane which can be related to the distance between
        pores. The feature, which will be called R-pores, is characterized by dots
        along $q_z$ = 0. The spacing between dots is indicative of the hexagonal
        symmetry of the packed pores. The same information at higher resolution is obtained using a SAXS
        setup. By radially integrating the 2D data one gets a 1D curve which is 
        shown in Figure~\ref{fig:SAXS}.  % BJC: I also have 2D SAXS data
  \end{itemize}

  \begin{figure}[!ht]
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
        %       \vspace{12mm}
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \raisebox{.2\textwidth}{%
                \includegraphics[width=\linewidth]{WAXS_soft_confined.png}
                }
                \caption{}\label{fig:WAXS}
        \end{subfigure}
        \caption{(a) 2D wide angle X-ray scattering gives details about repeating
        features less than 1 nanometer apart. (b) 1D small angle X-ray scattering
        indicates hexagonal packing of pores as well as the spacing between pores.}\label{fig:SWAXS}
  \end{figure}

%  We chose an initial layer spacing based on experimental 2D WAXS data.
%  \begin{itemize}  
%    \item The pattern shows reflections corresponding to features spaced 3.7 \angstrom~apart.
%    \item It has been hypothesized that the features are a result of $\pi$-$\pi$
%    interactions between stacked aromatic rings \cite{feng_scalable_2014}. 
%    \item Our simulations tend to equilibrate to a wider interlayer spacing of
%    $\approx$ 4.1 \angstrom~, which inspired separate systems starting with layer 
%    spacings greater than 4 \angstrom~.
%  \end{itemize}

  We chose the initial layer spacing based on R-$\pi$.
  \begin{itemize}  
    \item Each monomer was rotated so the plane of the aromatic head groups would
    be parallel to the xy plane.
    \item The layers are placed so aromatic rings are stacked 3.7 \angstrom~
    apart in the z-direction.
    \item To quantify the degree of layering in our system, we calculated
    a spatial correlation function, $g(z)$, measured along the z-axis 
    (perpendicular to the membrane plane)
    \item To calculate $g(z)$, we binned the z component distances between 
    the center of mass of each benzene ring and all others of the same pore 
    over 50 ns of equilibrated trajectory and then normalized by the average
    number density.
    \item We compare the degree of layering between systems based on the
    amplitude of the first peak in $g(z)$. We halve the difference between the
    maximum of the first peak and the following minimum. We compare the difference
    to the mean to get a percentage deviation from the average number density.
    \item To extract the average distance between layers we applied a discrete 
    fourier transform to $g(z)$ and extracted the highest intensity frequency 
  \end{itemize}

  We placed pores at a chosen initial spacing based on R-pores, then allowed
  the system to settle into its preferred spacing. 
  \begin{itemize}
    %MRS: somewhere, should explain there's a large dependence of (meta)stable structures based on this initial distance.  It's sort of important to put somewhere.
    %BJC: Sure, but different choices of pore spacing were based on number of monomers per layer. More monomers per layer means a larger initial pore spacing.
    \item The model's pore centers are spaced 4.5 nm apart initially, $\approx$ 10 \% larger
    than the experimental value of 4.12 nm in order to reduce unintended repulsions 
    resulting from a tightly packed initial configuration.
    \item To calculate the equilibrated pore spacing, we measured the distance between pore
    centers.
    \item Pore centers were located by averaging the coordinates of sodium ions in their 
    respective pores.
    \item Statistics were generated using the bootstrapping technique (See Supplemental Information)
    %BJC: Do bootstrapping details go in supplemental?
    	% Supplemental vvvv
	\begin{itemize}
    	\item For each bootstrap trial, we recreate an equilibrium trajectory by randomly 
    	sampling from the original trajectory
    	\item Each pore spacing has its own trajectory with its own average value
    	\item The average value of each pore spacing is averaged to get the overall
    	average pore spacing. 
    	\item The standard deviation of average values is reported as the uncertainty in pore spacing
    	% BJC: The following will go in supplemental
        \item We are interested in 5 pore-to-pore distances which should all be equal in 
        a perfect hexagonal array. Only 4 are independent  
        % Supplemental ^^^^
	\end{itemize}
  \end{itemize}

  %MRS: pore radius is a different category isn't it? Is this an input piece of data?
  % should be put in a different category.
  %BJC: Well I still have to initially place the monomers some distance away from the pore
  % centers. That distance is informed by the experimental evidence.

%  We based the pore radii of our initial configuration on past TEM and size exclusion 
%  rejection data
  We used experimental Transmission electron microscopy (TEM) and size exclusion rejection
  data \cite{feng_scalable_2014,feng_thin_2016,zhou_supported_2005} to inform our definition
  of pore radius in the initial configuration.
  \begin{itemize}
    \item Experimental evidence suggests uniform pores with radii of 0.6 nm 
    \item Comparing a geometric measurement of pore size derived from an atomistic model,
    to a less precise, experimentally derived pore size estimate, will give ambiguous results.
    \item What is meant by pore radius will not be clear until we establish a clear picture
    of the nanoscopic pore environment.
    \item When constructing pores, we chose the carboxylate carbon from the monomer
    head group as a reference atom, and placed it a distance r from the pore center,
    where r is the pore radius. (See Supplemental Information) %BJC: figure making that clear
    \item We will not make direct comparisons of pore radius between our model 
    and experiment to avoid the ambiguity, however, we do define a pore radius  
    for our own purposes.
    %\item To measure the pore radius in our model we calculate the distance between 
    %the center of mass of each aromatic ring and the center of mass of all aromatic rings
    %in their respective pores.
  \end{itemize}

  The relative interlayer orientation was chosen based on clues from 
  diffraction data as well as the various known stacking modes of benzene 
  and substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  ~\cite{sinnokrot_estimates_2002} (\Cref{fig:sandwiched,fig:pd,fig:tshaped}).
  \begin{itemize}
    \item The T-shaped configuration was ruled out based on the inconsistency of
    its $\approx$ 5 \angstrom~equilibrium stacking distance ~\cite{sinnokrot_estimates_2002}.
    \item The system's preference towards the sandwiched vs. parallel displaced 
    stacking modes will be explored.
    \item Both have reported stacking distances near the R-$\pi$ value of 3.7 \angstrom % Reference
    \item Headgroups in the sandwiched configuration are stacked directly on top of
    each other while stacked headgroups in the parallel displaced configuration are offset 
    by $180/nmon$ degrees where $nmon$ equals the number of monomers per layer.
    %BJC: I think the following commentary belongs elsewhere, maybe even in the discussion
    %MRS: hard to say. It's certainly a big switch from things that are relatively straightforward to things that require 
    % a fair amount of information, so I think it belongs in it's own section -- so far, you're just laying out the approach.  This is what we fix, this is what we leave as a variable.  The analysis probably belongs later.
    %BJC: Yeah, I will leave the rest for discussion
%     \item Visualization of each configuration (\Cref{fig:sandwichedlayers,fig:offsetlayers})
%     suggests entropic differences based on the way the tails are able to pack.
%     \item In the sandwiched configuration, all tails start out directly on top
%     of each other which may prevent closely stacked benzene rings.
%     \item In the offset configuration, the tails are placed in between each other 
%     which may allow layers to come together in a compact way.
%     \item This difference may explain, in part, which stacking mode is more favorable.
  \end{itemize}

  %We tested a number of equilibration schemes of increasing complexity in an effort
  %to match experiment while overcoming kinetic limitations that might trap the system 
  %in a metastable states. All start from an initial configuration generated with 
  %chosen structure variables.
  We developed equilibration schemes to create dry and wet configurations. Both schemes 
  start with an initial configuration generated according to the previous guidelines.
  \begin{itemize}
      \item Dry equilibration scheme:
      \begin{itemize}
          \item Restraints fix monomer head groups in the sandwiched or parallel-displaced
          configurations while allowing monomer tails to settle.
          \item Doing so also mitigates system dependence on initial monomer configuration.
          \item The restrained portion of the equilibration scheme is run in the NVT ensemble.
          \item Every 50 ps, we reduce the force constants by the square root of its
          previous value, starting from 1e6 KJ mol$^-1$ nm$^-2$.
          \item Once the force constant is below 10 KJ mol$^-1$ nm$^-2$, the restraints are
          slowly released until there is no more restraining potential.
          \item The resulting unrestrained structure is allowed to equilibrate further in the NPT
          ensemble for at least 400 ns.
      \end{itemize}
      \item Wet equilibration scheme: 
      \begin{itemize}
	  % BJC: the following three bullet points ended up not being the case since I used significantly less water
          %\item We obtained an estimate of equilibrium pore water content by solvating
          %an initial configuration with water baths above and below the membrane. 
	  %\item We ran the solvated system according to Scheme A but with a 1000 ns equilibration
          %\item Using that number we added water to the pores of an initial 
	  %configuration and equilibrated according to scheme B. 
	  \item In order to create a wet system, we solvated an initial configuration with water using gmx solvate
          \item All waters placed outside the pore region are removed 
	  \item Waters inside the pore region are randomly removed until the desired concentration
	  of water in the pores is reached. 
	  \item The remainder of the equilibration is the same as the dry system
      \end{itemize}
      \item In all cases, the v-rescale thermostat was used with tau-t = 0.1 % BJC: supplemental?
  \end{itemize}

%  An equilibration scheme with position restraints placed on aromatic rings
  %MRS: 
  %prevents 
%  during early equilibration steps is required to prevent high energy repulsions .
  %MRS: should be clearer what you mean by ``unrealistic jumps''.  Hard to visualize.
  %BJC: 'high energy repulsions' ?
%  \begin{itemize}
%    \item Restraints fix monomer head groups in the sandwiched or parallel-displaced
%    configurations while allowing monomer tails to settle.
%    \item Doing so also mitigates system dependence on initial monomer configuration.
%    \item The restrained portion of the equilibration scheme is run in the NVT ensemble.
%    \item Every 50 ps, we reduce the force constants by the square root of its 
%    previous value, starting from 1e6 KJ mol\textsuperscript{-1} nm\textsuperscript{-2}.
%    \item Once the force constant is below 10 KJ mol\textsuperscript{-1} 
%    nm\textsuperscript{-2}, the restraints are slowly released until there is no more 
%    restraining potential.  
%    \item The resulting unrestrained structure is allowed to equilibrate further in the NPT
%    ensemble for 400 - 500 ns.
   %BJC: I will reference an equilibration script that can be released with the supplemental info
   %MRS: we should release as many scripts as is feasible, can be in a github repository.
   %BJC: Okay, once we submit, I will need to spend some time cleaning the scripts
   %MRS: You've described two equilibration procedures (or two varieties of the same one?) - need to be crystal clear to the reader as to which is used where.
   %BJC: Yea I'm just saying that I do 50 ps restrained simulations in NVT ensemble and then once the restraints are gone I switch to NPT. I edited the language to try and make it clearer.
%  \end{itemize}

  %BJC: the following paragraph will be replaced with something that Joe writes
  %MRS: sounds good.

  % BJC: The following methodology is taken from the 2014 Feng et al. paper where
  % WAXS is reported. The data we are using comes from the 2016 paper which 
  % I assume used the same set up but I will need to verify this with Xunda
  % before this gets sent out. The SAXS is the same as shown in the 2016 paper
  % however I will likely reproduce it with matplotlib

  %Experimental wide angle X-ray scattering measurements were performed using a 
  %Rigaku 007 HF+ instrument with a rotating anode Cu K$\alpha$ X-ray source
  %($\lambda$~= 1.542 \angstrom~) and a 2-D Saturn 994+ CCD detector. The 
  %calibrations of the resultant 2-D WAXS patterns were done by using a 
  %silicon powder standard (d-spacing of 3.1355 \angstrom~). 2-D SAXS scattering
  %patterns were integrated into 1-D plots of scattering intensity (I) versus
  %q, where q = 4$\pi$sin($\theta$)/$\lambda$ and the scattering angles is
  %2$\theta$. The data shown is reproduced from data collected elsewhere~\ref{feng_scalable_2014}.

  % BJC: To be replaced by Joe's description
  Simulated X-ray diffraction patterns were generated based on atomic
  coordinates for a direct experimental comparison.
  \begin{itemize}
    \item All atomic coordinates were simulated as gaussian spheres of electron
    density corresponding to each atom's atomic number.
    \item A three dimensional fourier transform (FT) of the array of electron 
    density results in a three dimensional structure factor which represents
    the unit cell in reciprocal space.
   % \item We perform an angular average of the structure factor about the 
    z axis to generate a 2D cross section close to what one would see 
    experimentally.
    \item We matched experimental 2D WAXS patterns by iterative improvement
    of our choice of initial structure and equilibration procedure.
  \end{itemize}

>>>>

\documentclass{article}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[margin=1in]{geometry}

\begin{document}

\begin{figure}[ht]

\begin{subfigure}{.83\linewidth}
	\centering
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\linewidth, trim={2.5cm 0 4cm 1.25cm}, clip]{WAXS_raw.png}%
		\caption{}~\label{fig:raw_waxs}	
	\end{subfigure}%
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\linewidth, trim={2cm 0 2.5cm 1.25cm}, clip]{rzplot_offset.png}
		\caption{}~\label{fig:rz_offset}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\linewidth, trim={2cm 0 2.5cm 1.25cm}, clip]{rzplot_layered.png}
		\caption{}~\label{fig:rz_layered}
	\end{subfigure}%
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\linewidth, trim={2cm 0 2.5cm 1.25cm}, clip]{rzplot_disordered.png}
		\caption{}~\label{fig:rz_disordered}
	\end{subfigure}
\end{subfigure}%
\begin{subfigure}{0.14\linewidth}
	\centering
	\vspace{-1cm}
	\includegraphics[width=\linewidth]{colorbar_jet.png}
\end{subfigure}
\caption{(a) Experimental 2D WAXS data contains 5 major reflections which we aim to match. The remaining 
three images are diffraction patterns simulated from MD trajectories. (b) The parallel displaced
configuration gives rise to all reflections of interest. (c) The sandwiched configurations gives rise to a
pattern with all major reflections except R-helix. R-spots is strong relative to R-alkanes in comparison
to the parallel displaced configuration. (d) The disordered pore configuration creates a pattern with only
R-alkanes and R-pores in common with experiment}~\label{fig:xrd}
\end{figure}
